---
title: "Final Project Proposal"
author: "Suryansh Agrawal"
date: "11/19/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Team Members
<a href="http://suryanshagrawal.com/">Suryansh Agrawal</a> - Exploring Within

## Dataset 
<a href="https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-County/8xkx-amqh">COVID-19 Vaccinations in the United States</a>

## Investigation Questions
- How has the trend been? How is the US Government working towards the Global Pandemic (that still is not over)?
- Focusing down to the state where we live, Ohio and even Licking County, how has the status of COVID-19 changed?
- Can visualizing the data of the COVID-19 be helpful? Will visuals of even smaller areas help us improve COVID conditions? 

## Rationale Explanation
I found a data set from the US Government which is very much up to date and keeps on updating everyday. I found this dat set to answer a few questions that can, maybe, also be helpful in a larger extent. Having this data set can help me practice a lot of the skills that we learnt in this class. I see myself using  filtering, grouping, graphics, mapping, t-testing, and a lot more stuff that can be implemented on this data set. One thing that I am excited for the most is the use of maps and I feel that implementations of maps can be a super useful skill. Bringing down the co-ordinates to licking county and analyzing it can also give answers to the COVID situation near us and can help the students on campus. 

---

### A glace to the data set

```{r}
# Installing Library 
library(tidyverse)

# Reading the COVID data
covid <- read.csv("COVID-19_Vaccinations_in_the_United_States_County.csv")
summary(covid)

# Making a dataset of Ohio
ohio <- filter(covid, Recip_State == "OH")
summary(ohio)

# Making a dataset of Licking County
licking <- filter(ohio, Recip_County == 'Licking County')
summary(licking)

# The dataset
covid
```

